Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending

### 2017-07-25 paper.Rmd
big update

bibliography: library.bib
output: rticles::elsevier_article
---

@@

# Packages (load child-specific packages in their respective files)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(RColorBrewer)
# library(Hmisc)
# library(Rmisc)

@@

# Load data
file_name <- paste0('snapshot_', Sys.Date(), '.RData')
if(!file_name %in% dir('snapshots')){
  # Fetch data
  library(gsheet)
  url_responses<- 'https://docs.google.com/spreadsheets/d/1yeQ4TR73mdnDABPZ-ZEg3DAFu2Bko5NJ_nkMcNKr2lo/edit?usp=sharing'
  df <- gsheet2tbl(url_responses)

  # Manually rename variables
  names(df) <- tolower(gsub('.', '_', names(df), fixed = TRUE))
  names(df)[4:17] <-
    c('area',
      'years_10',
      'years_20',
      'years_30',
      'years_40',
      'years_50',
      'years',
      'governance',
      'health_workforce',
      'health_financing',
      'health_technologies',
      'health_information',
      'service_delivery',
      'comments')

  # Save a snapshot
  save(df, 
       file = paste0('snapshots/',
                     file_name))
} else {
  load(paste0('snapshots/',
              file_name))
}
```

```{r}
# Clean up survey responses
# Make NA all empty responses
df$comments[df$comments == ''] <- NA
df$email_address[df$email_address == ''] <- NA
# Flag duplicates
df$duplicate <- duplicated(df$email_address)
# Remove duplicates
df <- df[is.na(df$email_address) |
           !df$duplicate,]
# Create area dummies
df$area <- tolower(df$area)
df$area <- gsub('/', ', ', df$area)
df$area <- gsub('(', '', df$area, fixed = TRUE)
df$area <- gsub(')', '', df$area, fixed = TRUE)
# Perform some replacements
df$area[df$area == 'biochemistry'] <- 'chemistry'
df$area[df$area == 'medical entomology'] <- 'entomology'
df$area[df$area == 'vector biology'] <- 'biology'

@@

# Get number of areas
# Average number of areas
area_counter <- function(x){
  result <- unlist(gregexpr(',', x))
  if(result[1] == -1){
    return(1)
  } else {
    return(length(result) + 1)
  }
}
df$n_areas <- sapply(df$area, area_counter)

@@

# Clean up the rankings
bb <- c('governance', 'health_workforce',
        'health_financing', 'health_technologies',
        'health_information', 'service_delivery')
for (i in 1:nrow(df)){
  sub_df <- df[i,bb]
  new_values <- sub_df / sum(sub_df, na.rm = TRUE) * 21
  new_values[new_values > 6] <- 6
  df[i,bb] <- new_values
}

```{r}
# Results cleanup
```

@@

@@

@@

@@

>>>>

bibliography: library.bib
params:
  get_new_data: False
output: rticles::elsevier_article
---

@@

# Packages (load child-specific packages in their respective files)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(RColorBrewer)
# library(Hmisc)
# library(Rmisc)
# install.packages(c("gender", "genderdata"),
#                  repos = "http://packages.ropensci.org",
#                  type = "source")
library(gender)
library(wru)
library(sampleSelection)
library(stargazer)

@@

# Load data
get_new_data <- params$get_new_data
if(get_new_data){
  file_name <- paste0('snapshot_', Sys.Date(), '.RData')
} else {
  snaps <- dir('snapshots')
  dates <- as.Date(substr(snaps, 10, 19))
  file_name <- snaps[which.max(dates)]
}
if(!file_name %in% dir('snapshots')){
  # Fetch data
  library(gsheet)
  url_responses<- 'https://docs.google.com/spreadsheets/d/1yeQ4TR73mdnDABPZ-ZEg3DAFu2Bko5NJ_nkMcNKr2lo/edit?usp=sharing'
  df <- gsheet2tbl(url_responses)

  # Manually rename variables
  names(df) <- tolower(gsub('.', '_', names(df), fixed = TRUE))
  names(df) <- gsub(' ', '_', names(df))
  names(df)[4:17] <-
    c('area',
      'years_10',
      'years_20',
      'years_30',
      'years_40',
      'years_50',
      'years',
      'governance',
      'health_workforce',
      'health_financing',
      'health_technologies',
      'health_information',
      'service_delivery',
      'comments')

  # Save a snapshot
  save(df, 
       file = paste0('snapshots/',
                     file_name))
} else {
  load(paste0('snapshots/',
              file_name))
}

# Clean up survey responses
# Make NA all empty responses
df$comments <- ifelse(df$comments == '', 
                      NA,
                      df$comments)
df$email_address <- ifelse(df$email_address == '',
                           NA,
                           df$email_address)
# Flag duplicates
df$duplicate <- duplicated(df$email_address)
# Remove duplicates
df <- df[is.na(df$email_address) |
           !df$duplicate,]
# Create area dummies
df$area <- tolower(df$area)
df$area <- gsub('/', ', ', df$area)
df$area <- gsub('(', '', df$area, fixed = TRUE)
df$area <- gsub(')', '', df$area, fixed = TRUE)
# Perform some replacements
df$area[df$area == 'biochemistry'] <- 'chemistry'
df$area[df$area == 'medical entomology'] <- 'entomology'
df$area[df$area == 'vector biology'] <- 'biology'

@@

# Get number of areas
# Average number of areas
area_counter <- function(x){
  result <- unlist(gregexpr(',', x))
  if(is.na(result)){
    return(0)
  } else if(result[1] == -1){
    return(1)
  } else {
    return(length(result) + 1)
  }
}
df$n_areas <- sapply(df$area, area_counter)

@@

# Clean up the rankings
bb <- c('governance', 'health_workforce',
        'health_financing', 'health_technologies',
        'health_information', 'service_delivery')
for (i in 1:nrow(df)){
  sub_df <- df[i,bb]
  new_values <- sub_df / sum(sub_df, na.rm = TRUE) * 21
  new_values[new_values > 6] <- 6
  df[i,bb] <- new_values
}
```

```{r}
# Get the non-respondents into the same data
df$responded <- TRUE
non_respondents <- email_df %>%
  mutate(surname = `first _author_last_name`,
         first_name = `first _author_first_name`,
         name = author) %>%
  dplyr::select(name, surname, first_name, email) %>%
  filter(!email %in% df$email)

# Add to one dataframe
df <- bind_rows(df, non_respondents)
```

```{r}
# Predict the gender based on first name ####

# Extract first/last names
df$first_name <- tolower(unlist(lapply(strsplit(df$name, ' '), function(x){x[1]})))
df$last_name <- tolower(unlist(lapply(strsplit(df$name, ' '), function(x){x[length(x)]})))
df$surname <- df$last_name

if(get_new_data){
  # Get gender based on name
  names_df <- df %>%
    group_by(first_name) %>%
    tally %>%
    ungroup %>%
    dplyr::select(-n)
  # Get ethnicity based on last name
  surnames_df <- df %>%
    group_by(surname) %>% 
    tally %>%
    ungroup %>%
    dplyr::select(-n)
  gender_scores <- gender(names = names_df$first_name,
              years = c(1940, 1990),
              method = 'ssa') %>%
    rename(first_name = name) %>%
    dplyr::select(first_name,
                  proportion_male,
                  proportion_female,
                  gender)
  surname_scores <- 
    predict_race(voter.file = surnames_df,
                 surname.only = TRUE)
  save(surname_scores,
       gender_scores,
       file = 'data/name_scores.RData')
} else {
  load('data/name_scores.RData')
}

# Get most likely ethnicity
ethnicities <- c('white', 'black', 'hispanic',
                 'asian', 'other')
surname_scores$ethnicity <- ethnicities[apply(surname_scores[,2:6], 1, function(x){which.max(x)[1]})]

# Join surname scores to df
df <-
  left_join(x = df,
            y = surname_scores,
            by = 'surname')
df <-
  left_join(x = df,
            y = gender_scores,
            by = 'first_name')

# Flag non respondents as such
df$responded[is.na(df$responded)] <- FALSE
```

```{r}
# sentiment score of contents
library(databrew)
df$sentiment <- databrew::score_sentiment(df$comments)
```

```{r}
# Heckman selection
# We're estimating years as a function of sex and ethnicity
ols1 = lm(years ~ sentiment  + area_epidemiology + area_anthropology, data = df[df$responded,])

# Estimate likelihood of response
# it is very desirable to have at least one variable in the selection equation that acts like an instrument:
# (https://rpubs.com/wsundstrom/t_selection)
# for us, this is ethnicity - it should affect response (due to language)
# but not likely to affect optimism

# Heckman 2 step
heck1 <- heckit(responded ~ gender + ethnicity,
                years ~ sentiment + area_epidemiology + area_anthropology, data = df)

# Maximum likelihood (like heckman 2 step, but slightly different)
# ML estimation of selection model
ml1 = selection(responded ~ gender + ethnicity,
                years ~ sentiment + area_epidemiology + area_anthropology, data = df)

# # Look at results
# stargazer(ols1, heck1, ml1,    
#           # se=list(cse(ols1),NULL,NULL), 
#           title="Regression on years until eradication", type="latex", 
#           df=FALSE, digits=4)
```

@@

## Data generation

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers' demographics is of value. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name, and the likelihood of a participant being of one of five ethnicities (White, Black, Hispanic, Asian, Other) based on the provided last name.

@@

The estimation of gender and ethnicity likelihoods for each participant relied on Mullen's implementation of US Social Security Administration names' analysis [@gender2015] and Khanna et al's bayesian prediction algorithm for the estimation of racial categories [@wru2015].

De-biasing of sample selection was carried out using Heckman's 2 step approach (citation needed), as implemented by Toomet [@toomet2008].

@@

## Sample selection and de-biasing of results

This section is unfinished. Below are regression outputs for both OLS and Heckman's 2 step model. I will update this soon.

```{r}
# stargazer(ols1, heck1, ml1,    
#           # se=list(cse(ols1),NULL,NULL), 
#           title="Regression on years until eradication", type="html",
#           df=FALSE, digits=4)
summary(ols1)
summary(heck1)

@@

## Sentiment analysis of free-text commentary

This section is unfinished. Below is the distribution of "sentiment" among the responses (0 is neutral, below 0 is negative, above 0 is positive), but this is preliminary. The chart is colored by ethnicity, and faceted by gender.

```{r}
ggplot(data = df %>%
         filter(responded) %>%
         # mutate(n_areas = factor(n_areas)) %>%
         filter(!is.na(ethnicity),
                !is.na(gender)),
       aes(x = sentiment,
           group = ethnicity)) +
  geom_density(aes(fill = ethnicity),
               alpha = 0.6,
               n = 10) +
  theme_publication() +
  facet_wrap(~gender)
